---
title: "Untitled"
author: "Arthur Harnisch"
date: "September 27, 2017"
output: html_document
---

#automating data analysis. writing your own functions in R. writing your own function 
# function name (argument = df true/false) when you write your own function you will be putting in your own arguments in .
# using the function function 

#the variable my function 

#function function ```{r functionExample}
#my function 
# curly brace defines start and end of funciton, additin line first multiplication line first 

```{r functionExample}
myfunction <- function (x){
  print(x)
}

```{r functionExample2}
myfunction (x=2)
myFunction (x= c(1,2,3,4)}

```{r functionExample 3)
myfunction2 <- function (x,y) {
  z= x+y
  z
}





```{r seFunction}
mySE <- function (x){
  #assumes the input x is a vector of numbers
  # first calculate the variance
  myvar = var(x)
  #then calculate the number of values
  myn = length(x)
  #next calculate the square root of the variance divided by sample size
  mySE = sqrt(myvar/myn)
  #hte value returned will be the last object created in the function, in this case mySE
  mySE #repeating just the last variable here will cause the result to print out when run
}
```

Now we can test that function
```{r testSE}

x=c(1,1,3,3,6,6)
mySE(x) #result should be 0.918..
```


avoid copy and paste. 
there are functions in R that will sometimes only take the name of another function as an argument. One example we've function. mecause we have created a function. there are 

in your analysis if you have multiple funnctions in the analysis you can stick in .rmd and source the .r script file of the function. 

```{r aggregateCustomFunction}
#make a practice data frame # note rep function means repeat. # this is a quick 
mydf = data.frame(group = c(rep(`A,`6) rep(`8,`6)),
                  values c (1,2,3,4,5,6)
mydf

mySE(mydf$Values[7:12])


aggregate(Values~group, data = mydf, FUN = mySE)


we now use our custom mySE function inside aggregate() to accomplish the same thing. 

```

hw you will create two functions download data and plot data and will do this several times

