---
title: 'Tide Plotting Homework 6'
author: "Arthur Harnisch"
date: "October 4, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
We will download and plot several months of tide height data for a given NOAA tide station. To accomplish this, we will use a function tideDownload designed to access the NOAA tide data website and request 1 month of data at a time. The website limits the user to a maximum of 1 month per request.

First we will request data for Monterey CA (station 9413450) for the month of January 2017.In order to do this we are going to create a new funciton tideDownload that will get the infomration from the NOAA tide station. 



```{r tideFunction}

################################################################################
# Run this script as-is. It will prompt you to enter the appropriate
# station and date values when needed. You can only request less than 1 month of 
# data at a time, so if you get an error returned, try asking for slightly less
# data.

# This script will download a set of verified tide height data from a NOAA 
# CO-OPS DODS/OPeNDAP server and parse it into a data frame to be saved to disk 
#
# Station list: http://opendap.co-ops.nos.noaa.gov/stations/index.jsp
# Scroll through that list to find your station ID. 

# OPeNDAP server gateway: https://opendap.co-ops.nos.noaa.gov/erddap/index.html
# The gateway has links to several other data types available including 
# water temperature, air temperature, wind etc. 
# See http://docs.opendap.org/index.php/UserGuideOPeNDAPMessages for info on
# structuring OPeNDAP queries.
#
# Six-minute water level data can only be retrieved 1 month at a time. Other
# fields such as water temperature could return up to 1 year at a time. This 
# script only deals with 6-minute tide height data, but could serve as a guide 
# for accessing other data types from NOAA CO-OPS. 
# The tide height is reported in meters and time in UTC (Greenwich).

###############################################################################
library(RCurl)


tideDownload <- function(station, startdate, enddate){




#noquote(print("Enter NOAA tide station number (i.e. Monterey = 9413450): "))
#station = scan("",what = character(),nlines = 1)
#noquote(print("Enter start date (format: 20080123 = Jan 23, 2008): "))
#startdate = scan("",what = character(),nlines = 1) #get one line of values
#noquote(print("Enter end date (format: 20081231 = Dec 31, 2008): "))
#enddate = scan("",what = character(),nlines = 1)

#OPeNDAP query for 6-minute verified water level looks like this (on 1 line):
# https://opendap.co-ops.nos.noaa.gov/erddap/tabledap/
# IOOS_SixMin_Verified_Water_Level.asc?STATION_ID
# %2CDATUM%2CBEGIN_DATE%2CEND_DATE%2Ctime%2CWL_VALUE%2CSIGMA%2CI%2CF%2CR%2CT&
# STATION_ID=%229413450%22&
# DATUM=%22MLLW%22&
# BEGIN_DATE=%2220080123%22&
# END_DATE=%2220080130%22
#####################################################
######## DON'T CHANGE ANYTHING BELOW THIS LINE ####################
#The parts of the url to be assembled:
url1 = "https://opendap.co-ops.nos.noaa.gov/erddap/tabledap/"
url2 = "IOOS_SixMin_Verified_Water_Level.asc?"
url3 = "STATION_ID%2C" #return stationId
url4 = "DATUM%2C" #return datum
url5 = "time%2C" #return record date-time
url6 = "WL_VALUE%2C" #return water level value
url7 = "I%2C" #return quality flag
url8 = "F%2C" #return quality flag
url9 = "R%2C" #return quality flag
url10 = "T" #return quality flag
#The remaining parts of the url specify how to filter the data on the server 
#to only retrieve the desired station and date range. Values must be enclosed
#in ascii double-quotes, which are represented by the code %22
# Do not change any values here, do not insert your own values here.
url11 = "&STATION_ID=%22" # station ID goes here
url12 = "%22"
url13 = "&DATUM=%22MLLW%22" # we want MLLW as the datum
url14 = "&BEGIN_DATE=%22" # start date gets put in here
url15 = "%22"
url16 = "&END_DATE=%22" # end date gets put in here
url17 = "%22"
####### DON'T CHANGE ANYTHING ABOVE THIS LINE ############
##############################################

#Assemble the URL
urltotal = paste(url1,url2,url3,url4,url5,url6,url7,url8,url9,url10,url11,
		station,url12,url13,url14,startdate,url15,url16,enddate,url17,sep ="")

#Download the data
cat("Contacting server...\n"); flush.console()
dat = getURL(urltotal) #use RCurl to retrieve text into a vector 'dat'
cat("Data returned...\n"); flush.console()
Sys.sleep(2) #If you access data in a loop, be courteous and give the server
#a short break between requests
#cleanup
rm(url1,url2,url3,url4,url5,url6,url7,url8,url9,url10,url11,url12,url13,url14)
rm(url15,url16,url17)

con = textConnection(dat) #create text Connection to dat vector
all.lines = readLines(con) #read lines of text into separate slots in a vector
close(con) #close connection to dat vector

if (length(grep('^Error',all.lines))>0) { #check for error in retrieval
	cat("There was an error...\n")
	cat(dat,"\n") #print contents of dat to show error
	flush.console()
} else { #retrieval was successful, parse the text
	
	#The column headers are typically preceded by a line of dashes
	headerlines = grep("^--------",all.lines) #find index of headers (-1)
	
	#read column header names into a vector
	con = textConnection(dat)
	headers = scan(con, skip = headerlines, nlines = 1, sep = ",",
			what = "character", strip.white = TRUE)
	close(con)
	
	#read rest of the data into a data frame 'df'
	con = textConnection(dat)
	df = read.table(con, skip = headerlines+1, sep = ",", header = FALSE,
			quote = "\"", col.names = headers, strip.white = TRUE,
			stringsAsFactors = FALSE)
	close(con)
	
#The following operations will need to be altered if you change the 
	#fields or data type being returned by the OPeNDAP server
	# The time column should be a numeric value, representing elapsed 
	# seconds since midnight, Jan 1, 1970 in the GMT (UTC) time zone. 
	df[,3] = as.POSIXct(df[,3], tz = 'GMT', origin = '1970-1-1 00:00')
	
	#Give the columns shorter names
	names(df) = c("stationId","datum","TimeUTC","TideHT","Flag.Inferred",
			"Flag.Flat.Tol","Flag.Rate.Tol","Flag.Temp.Tol")
	
	#Uncomment this if you want to plot the data
	#plot(df$TimeUTC, df$TideHT, type = "l",
			#xlab = "Date",ylab = "Tide Height, meters")
	
	#Save data automatically to a .csv file. 
#	filename = paste("Station_",station,"_",startdate,"-",enddate,
	#		".csv",sep = "") #make file name
	#write.csv(df,filename,row.names = FALSE, quote = FALSE) #write file to disk
	#cat("Saved to ",getwd(),"/",filename,"\n",sep = "")
	#flush.console()
	
	#Alternate file save method lets user specify file name at run time
	#Uncomment this if you wish to use it instead of the automated file 
	#output above
#	write.csv(df,file.choose(),row.names = FALSE, quote = FALSE)
	
	#cleanup
	rm(dat,con,all.lines,startdate,enddate,headerlines, headers)
	
	df
	
}} #end of if-else statement
```

``````{r plot function}


#create a plotting function for the tide data. 

plotTide <- function(df){
  plot(x = df$TimeUTC, y = df$TideHT, type = 'l',
       xlab = 'Date',
       ylab = 'Tide Height, m',
       col = 'blue')
}



```


We will now use the plotting code to make the plots for each month. The plot command could be wrapped up into its own function as well, which just takes the data frame tide heights as an argument. 


```{r grabJan, cache=TRUE, message = FALSE, warning = FALSE}
# using tideDownload function to get data

Jan2017= tideDownload(station = 9413450, startdate = 20170101, enddate = 20170131)

plotTide(Jan2017)
```


for the following months, we can repeat the same function using tideDownload to supply the new data range and the plotTide

```{r grabFeb, cache=TRUE, message = FALSE, warning = FALSE}


Feb2017= tideDownload(station = 9413450, startdate = 20170201, enddate = 20170228)

plotTide(Feb2017)
```


```{r grabMar, cache=TRUE, message = FALSE, warning = FALSE}


Mar2017= tideDownload(station = 9413450, startdate = 20170301, enddate = 20170331)

plotTide(Mar2017)
```







