---
title: "Homework Week 7"
author: "Arthur Harnisch"
date: "October 11, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing data

```{r importing data}

df = read.csv('../data/california-state-university-2015.csv')

csu = df$Employee.Name = as.character(df$Employee.Name)
```


Find all entries with the name "Martinez"

```{r finding Martinez}

#finding the number of people with Martinez in their name
length(grep("Martinez", df$Employee.Name, value = TRUE))

#listing the names that contain Martinez 
head(grep("Martinez", df$Employee.Name, value = TRUE))


```
Everyone whose name starts with the letters "Wil"
```{r finding Wil}

#finding the number of entries that begin with "Wil"
length(grep("^Wil", df$Employee.Name, value = TRUE))


#listing the individuals with name that begins with "Wil"
head(grep("^Wil", df$Employee.Name, value = TRUE))


```


Everyone with a hyphenated name

```{r finding hyphens}

#finding the number of people with hyphenated names
length(grep("-", df$Employee.Name, value = TRUE))

#listing the individual's with hyphenated names

head(grep("-", df$Employee.Name, value = TRUE))

```


Everyone whose middle name is paul

``` {r middleName}


# the majority of names on this list have an initial for the middle name for example

length(grep("\\sP\\s", df$Employee.Name, value = TRUE))
head(grep("\\sP\\s", df$Employee.Name, value = TRUE)) # this returns middle name begining with P


# However some people do have Paul as a middle name on this list though

length(grep("\\sPaul\\s", df$Employee.Name, value = TRUE))

head(grep("\\sPaul\\s", df$Employee.Name, value = TRUE))

```


All people with the last name Martinez

```{r lastNames}

length(grep("Martinez$", df$Employee.Name, value = TRUE))

head(grep("Martinez$", df$Employee.Name, value = TRUE))

```

All people with last name martinez but no hyphens or Jr.

```{r martinezAndnonMatching}


length(grep("(Martinez$)(?!-\\[jJ]r.)", df$Employee.Name, value = TRUE, perl = TRUE))


head(grep("(Martinez$)(?!-\\[jJ]r.)", df$Employee.Name, value = TRUE, perl = TRUE))


```

 All people with first name that starts with E has a middle initial and last name is Martinez

```{r specificMartinez}

#one way to do this is by using data frames 

eName=grep("^E", df$Employee.Name, value = TRUE)
mArtinez = grep("Martinez$", eName, value = TRUE)
length(grep("\\s.\\s", mArtinez, value = TRUE))
head(grep("\\s.\\s", mArtinez, value = TRUE))



#alternatevely this can also be accomplished by using 

length(grep("(^E)(.*\\s.\\s)(.*?Martinez$)", df$Employee.Name, perl = TRUE, value = TRUE))
grep("(^E)(.*\\s.\\s)(.*?Martinez$)", df$Employee.Name, perl = TRUE, value = TRUE)


```
First name starts with E has no middle initial and last name is martinez. 
```{r nonMiddle}

grep("(^E)(?!.*\\s.\\s)(.*?Martinez$)", df$Employee.Name, value = TRUE, perl = TRUE)


```




