---
title: "extra from HW 8"
author: "Arthur Harnisch"
date: "October 25, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:\\Users\\Arthur\\255E")
```

```{r extra}



```{r secondPlot}
heatingInfo = read.xlsx("../data/Heating_trial_info.xlsx", sheetIndex = 1)
heating1 = heatingInfo

sub1 = subset(heatingInfo, heatingInfo$TCLocation == "In50")

levels(sub1$TCchannel) <- sub('Ch', 'TC', levels(sub1$TCchannel))



if (heatingInfo$TCLocation[x] == "In50"){
  chloc = regexpr(pattern = "[[:digit:]]", hearingInfo$TCchannel[x])
  chNum = substr(heatingInfo$TCchannel[x], 
                  start = chloc[[1]][1], stop = chloc[[1]][1])
     chName = paste("TC", chNum, sep = "")
   trailNumber = heatingInfo$TrialNumber[x]
   
   temps = res[ res$trialNumber == trailNumber , chName]
   
  times= res$elapsedTimeMin[res$trialNumbe == trialNumber]
  
}




 for(x in 1:nrow(i)){
  
 if  (heatingInfo$TCLocation[x] == "In50") {
   chloc = regexpr(pattern = "[[:digit:]]", 
                   heatingInfo$TCchannel[x]) 
   chNum = substr(heatingInfo$TCchannel[x], 
                  start = chloc[[1]][1], stop = chloc[[1]][1])
   chName = paste("TC", chNum, sep = "")
   trailNumber = heatingInfo$TrialNumber[x]
   
   temps = res[ res$trialNumber == trailNumber , chName]
   
  times= res$elapsedTimeMin[res$trialNumbe == trialNumber]

   
  print(x)






#TC0~elapsedTime
#data = res[res$trialNumber ==1
#ploting trial 1 use plot command then lines command 

firstTrial =1
plot(x=temps[res$TrialNumber ==1], y= res$TC0[res$TrialNumber == firstTrial],
      
     type = 'l', 
     xlab = 'Elapsed time, Minutes',
     ylab = 'temperature C', las = 1,
xlim  = c(0,max(res$elapsedTime)),
ylim = c(min(res$TC0), max(res$TC0)))

mtext(text = "A", side = 3, line =1.4, cex =  1.5, 
adj = 0)

# elapsedTime)),


trialNums = unique(res$TrialNumber)

for(y in 2:length(trialNums) ){
  lines(x = res$elapsedTimeMin[res$TrialNumber == trialNums[y]],
  y=res$TC0[res$TrialNumber == trialNums[y]], col = y)
}




```
#plot(TC0~elapsedTimeMin, data = res, type = 'l')









plot(res3$TC0,res3$DateTime, type = "n")

points(res3$TC0, res3$DateTime, type = "p", pch=3 )

for(i in unique(res3$)) {
lines(res3[subj==i],time[subj==i],type="l",lty=2)
}














for(i in unique(subj)) {

lines(trial[subj==i],time[subj==i],type="l",lty=2) 


}






```


res$DateTime = as.POSIXct(res$DateTime, origin = '1970-1-1', tz="")



fnames = dir("./data/temperature_trials", 
             pattern = "*.csv", full.names = TRUE)

temp = list.files(pattern="*.csv")
myfiles = lapply(temp, read.delim)


```