---
title: "week 10"
author: "Arthur Harnisch"
date: "November 1, 2017"
output: word_document:
fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE), #fig.path = 'figs/', cache = TRUE # save copies of ifgures into folders

```


files get saved when knitting the whole document 

```{r pressurData, echo=TRUE}

summary(pressure)
pressure = pressure

```


adding dpi = 300 will increase the file dimensions


```{r pressure2, dpi = 300}

plot(pressure, col = 'blue')
```


```{r pressure3, dpi= 300, figwidth = 4, fig.height=7}

plot(pressure, col = red)
```

dev= c('png','pdf') c argument fits them together


saving figures
functions to save the fiel type. 

?png


```{r basicplorScript.eval = FALSE, echo= TRUE}

#dev.off #many lines of plotting codes will show up in image file but thi command is not necessary. 

```

1024/768 projectors

1920/1080 - projector screen across lab




###ggplot2 


```{r bacisSetup, echo=TRUE}

library(ggplot2)
diamonds = diamonds


```


```{rlayoutBasicPlot, fig.hsow='hide',echo=TRUE}

ggplot(diamonds)
ggplot(diamonds, aes(x=carat, y=price))

ggplot(diamonds, aes(x=carat, y=price))
```

```{r catatvscost, echo=TRUE, dpi=300}

ggplot(diamonds, aes(x=carat, y=price)) +
  geom_point()
```


```{redivdeByCUt, ECHO=TRUE, dpi=300}
#makes asthetic arguments 
ggplot(diamonds, aes(x=carat, y=price, color = cut)) +
geom_point()
```

```{rsmoothers, echo=TRUE, dpi= 300}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point(aes(color=cut)) + geom_smooth()


```{r singleSmooth, echo=TRUE, dpi=300}

ggplot(diamonds, aes(x=carat, y=price)) +
geom_point()+geom_smooth()

```




```{r shape}

ggplot(diamonds, aes(x=carat, y=price, color = cut, shape = color)) +
geom_point()
```


```{r ModColor, fig.cap='Transparency added'}

ggplot(diamonds, aes(x=carat, y=price, color= cut)) + geom_point()+ labs(title = "scatterplot", x = "Carat", y = "price")
```


```{r ST}

gg<-ggplot(diamonds, aes(x=carat, y=price, color= cut)) + geom_point()


gg= labs(title = "scatterplot", x = "Carat", y = "price")

print(gg)


```





```{r stuff}

gg1 = gg +theme(plot.title=element_text(size=30, face = "bold"), 
axis.text.x = element_text(size = 18), 
axis.text.y = element_text(sizex=18), 
axis.title.x=elements_text(size=25), 
axis.title.y=element_text(size = 25)))
```

```{r facet1, echo=true, dpi =300}
gg+facet_wrap( ~ cut, ncol = 3) 

# allows for multiple plots for one group 
```

```{r barchart1, echo=TRUE, dpi=300}

mtcars= mtcars
ggplot(mtcars, aes(x=cyl)) + geom_bar() + labs(title = 'Frequency bar chart')

```


```{r boxplot}
ggplot(mtcars, aes(x=factor(cyl), y=hp)) + stat_boxplot()

```



```{r volinplot}

ggplot(mtcars, aes(x=factor(cyl), y=hp)) + geom_violin()

```

```{r boxplot}

ggplot(mtcars, aes(x=factor(cyl), y=hp)) +geom_jitter(width = 0.1, height = 0, aes(color = factor(cyl)))

```
```